---
title: "Models of Neighborhood Change in Dallas - Fort Worth, TX"
output: 
  flexdashboard::flex_dashboard:
    theme: cerulean
    source: embed
    smart: false
runtime: shiny
---

    

```{r global, echo=F}

# PACKAGES 


# dashboard layout, widgets, and uploading

library( flexdashboard )
library( shiny )
library( rsconnect )

# data wrangling 

library( dplyr )


# formatting output

library( DT )
library( pander )
library( knitr )
library( stargazer )


# maps

library( ggmap )
library( leaflet )
library( viridis )
library( geojsonio )
library( sp )
library( sf )
library( tmap )
library( pals )
library( rgdal )

```



```{r, include=FALSE}

# DATA STEPS 

# load dorling cartogram from github
# map already contains census data and groups from clustering 

github.url <- "https://github.com/RTrehern-ASU/NHood-Change/blob/main/DFW2.geojson"
DFW <- geojson_read( x=github.url,  what="sp" )

# reproject the map 
DFW <- spTransform( DFW, CRS("+init=epsg:3395") )

# convert the sp map format to 
# an sf (simple features) format:
# ggmap requires the sf format
DFW.sf <- st_as_sf( DFW )

# separate out the data frame from the map
d <- as.data.frame( DFW.sf )
```


Community Demographics  
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
dem.variables <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12",
                     "polang12", "phs12", "pcol12", "punemp12", "pflabf12",
                     "pprof12", "pmanuf12", "pvet12", "psemp12", "hinc12",
                     "incpc12", "ppov12", "pown12", "pvac12", "pmulti12",
                     "mrent12", "mhmval12.x", "p30old12", "p10yrs12", "p18und12",
                     "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

dem.names <- c("Percent white, non-Hispanic", "Percent black, non-Hispanic",
                "Percent Hispanic", "Percent Native American race",
                "Percent foreign born", "Percent speaking other language at home, age 5 plus",
                "Percent with high school degree or less", "Percent with 4-year college degree or more",
                "Percent unemployed", "Percent female labor force participation",
                "Percent professional employees", "Percent manufacturing employees",
                "Percent veteran", "Percent self-employed", "Median HH income, total",
                "Per capita income", "Percent in poverty, total", "Percent owner-occupied units",
                "Percent vacant units", "Percent multi-family units", "Median rent",
                "Median home value", "Percent structures more than 30 years old",
                "Percent HH in neighborhood 10 years or less", "Percent 17 and under, total",
                "Percent 60 and older, total", "Percent 75 and older, total",
                "Percent currently married, not separated", "Percent widowed, divorced and separated",
                "Percent female-headed families with children")

radioButtons( inputId="demographics", 
              label = h3("Census Variables"),
              choiceNames=dem.names,
              choiceValues=dem.variables,
              selected="pnhwht12")
```


Row {.tabset}
-------------------------------------


### Choropleth Map


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             DFW.sf <- 
             DFW.sf %>% 
             mutate( q = ntile( get(input$demographics), 10 ) )  
          })

ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Choropleth of Select Demographics: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Population Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -10900000, xmax = -10700000 ) + 
    ylim( ymin = 3790000, ymax = 3920000 )
})
```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()
  
get_variable_x <- reactive({ d[ input$demographics ] })

x <- get_variable_x() %>% unlist()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of variable ", toupper( input$demographics ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )
})
```




Neighborhoods  
===================================== 

### Clusters  

```{r}

# define the bounding box corners 

bb <- st_bbox( c( xmin =  -10900000, xmax = -10700000, 
                  ymax = 3920000, ymin = 3790000 ), 
               crs = st_crs("+init=epsg:3395"))

# ADD YOUR CUSTOM LABELS TO THE CLUSTERS 

DFW$cluster[ DFW$cluster == "1" ] <- "Higher % black, veterans, 60+"
DFW$cluster[ DFW$cluster == "2" ] <- "Higher % white, high per-capita income, 60+"
DFW$cluster[ DFW$cluster == "3" ] <- "Higher % married homeowners, higher median HH income"
DFW$cluster[ DFW$cluster == "4" ] <- "Higher % college grads, prof employees w/higher MHV"
DFW$cluster[ DFW$cluster == "5" ] <- "Higher % veterans 75+, higher % female labor force"
DFW$cluster[ DFW$cluster == "6" ] <- "Higher % Hispanic w/less than high school degree"
DFW$cluster[ DFW$cluster == "7" ] <- "Higher % white w/col degree and higher MHV"
DFW$cluster[ DFW$cluster == "8" ] <- "Higher % multi-fam units and female labor force"
DFW$cluster[ DFW$cluster == "9" ] <- "Higher % Hispanic, female headed fam w/children"


renderTmap({

tmap_mode("view")
tm_basemap( "CartoDB.Positron" )
tm_shape( DFW, bbox=bb) + 
  tm_polygons( col="cluster", palette="Accent",
               title="Community Types", legend.position = c("left", "bottom"))
})
```




NH Change 2000-2010  
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}
button.labels <- c("Median Home Value 2000","Median Home Value 2010",
                   "Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.00","mhv.10","mhv.change","mhv.growth")
  
radioButtons( inputId="home.value", 
              label = h3("Home Values"),
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.10")
```




Row {.tabset}
-------------------------------------



### Median Home Values


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             DFW.sf <- 
             DFW.sf %>% 
             mutate( q = ntile( get(input$home.value), 10 ) )  
          })

  
ggplot( get_data(), na.rm=T ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Spatial Distribution of Home Values: ", toupper(input$home.value) ),
          caption = "Source: Harmonized Census Files",
          fill = "Home Value Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -10900000, xmax = -10700000 ) + 
    ylim( ymin = 3790000, ymax = 3920000 )
})

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$home.value ] })

x <- get_variable_x() %>% unlist() %>% as.numeric()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of ", toupper( input$home.value ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```  



Drivers of Change   
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.00","mhv.10","mhv.change","mhv.growth")
  
radioButtons( inputId="dv", 
              label = h3("Select Your Dependent Variable"),
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.00")


covariates <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12.x", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")


covariate.labels <- c("Percent white, non-Hispanic", "Percent black, non-Hispanic",
                "Percent Hispanic", "Percent Native American race",
                "Percent foreign born", "Percent speaking other language at home, age 5 plus",
                "Percent with high school degree or less", "Percent with 4-year college degree or more",
                "Percent unemployed", "Percent female labor force participation",
                "Percent professional employees", "Percent manufacturing employees",
                "Percent veteran", "Percent self-employed", "Median HH income, total",
                "Per capita income", "Percent in poverty, total", "Percent owner-occupied units",
                "Percent vacant units", "Percent multi-family units", "Median rent",
                "Median home value", "Percent structures more than 30 years old",
                "Percent HH in neighborhood 10 years or less", "Percent 17 and under, total",
                "Percent 60 and older, total", "Percent 75 and older, total",
                "Percent currently married, not separated", "Percent widowed, divorced and separated",
                "Percent female-headed families with children")


checkboxGroupInput( inputId="covariates", 
              label = h3("Select Variables for Your Model"),
              choiceNames=covariate.labels,
              choiceValues=covariates,
              selected=c("pnhwht12","pprof12","pvac12") )

```




Row {.tabset}
-------------------------------------



### Predicting Change 

```{r, results="asis"}

get_covariates <-  reactive({ input$covariates })
get_dv <-  reactive({ input$dv })

renderUI({
  
covariates <- get_covariates() 
dv <- get_dv()

formula.text <- paste0( dv, " ~ ", paste( covariates, collapse=" + " ) )
fo <- as.formula( formula.text )

m <- lm( fo, data=d )


# HTML( "<div style='width: 60%; margin: 0px auto;'>" )
HTML( 
  
  c("<br><br><br>",
  "<div type='regression' style='width: 60%; margin: 0px auto;'>",
  stargazer( m, type="html", omit.stat=c("rsq","f") ),
  "</div>",
  "<br><br><br>"
  )
  
  )
})

# HTML( reg.table )
```


<style>

.chart-shim { overflow: auto; }
 
table{
   border-spacing:1px;
   margin-top:30px;
   margin-bottom:30px;
   margin-left: auto;
   margin-right: auto;
   align:center} 

td{ padding: 6px 10px 6px 10px } 

th{ text-align: left; } 

</style>
